# tidyverse

```{r message=FALSE}
library(ggplot2)
```



- **factors** aims to handle categorical variables, i.e., `factor` class in R.



## `forcats::fct_reorder(f, x)`   

each level is associated with a `x` value (if multiple, use a function such as `median()` to generate a single value), the order is determined by sorting the corresponding `x` value.

```{r fct-reorder, collapse=TRUE}
with(iris, forcats::fct_reorder(Species, Sepal.Width)) %>% levels

iris %>% dplyr::group_by(Species) %>% 
   dplyr::summarise(Sepal.Width = median(Sepal.Width)) %>%
   dplyr::arrange(Sepal.Width) %>% .$Species %>% as.character()
```

As you can see in following plot, `Species` has been reordered by median of `Sepal.Width`.

```{r fct-reorder-plot, fig.cap='forcats::fct_reorder()'}
cowplot::plot_grid(
    iris %>% 
        ggplot(aes(Species, Sepal.Width)) + geom_boxplot() + 
        labs(title = 'original'),
    iris %>% dplyr::mutate_at("Species", ~forcats::fct_reorder(., Sepal.Width)) %>%
        ggplot(aes(Species, Sepal.Width)) + geom_boxplot() + 
        labs(title = 'reorder')
)
```
    


## `forcats::fct_reorder2(f, x, y)`  

Each level is associated with many `x` values and many `y` values, we use a function such as `max()` to select a `x` value, then we can get the corresponding `y` value. Now each level is associated with a single `y` value, and we can determine the order by sorting those `y` values. [^fct-reorder2-plot]

```{r fct-reorder2}
set.seed(0)
chkw <- ChickWeight %>% head(60) %>% dplyr::mutate_at('Chick', forcats::fct_shuffle)

with(chkw, forcats::fct_reorder2(Chick, Time, weight)) %>% levels()

chkw %>% dplyr::group_by(Chick) %>% dplyr::arrange(dplyr::desc(Time)) %>% 
   dplyr::slice(1) %>% dplyr::ungroup() %>% 
   dplyr::arrange(dplyr::desc(weight)) %>% .$Chick %>% as.character()
```

As you can see in following plot, the vertical line (max `Time`), `Chick` has been reordered by `weight` at that line.

```{r fct-reorder2-plot, fig.cap='forcats::fct_reorder2()'}
cowplot::plot_grid(
    chkw %>% 
        ggplot(aes(Time, weight, colour = Chick)) + geom_point() + geom_line() + 
        geom_vline(xintercept = 21, alpha = 0.2) + labs(title = 'original'),
    chkw %>% dplyr::mutate_at("Chick", ~forcats::fct_reorder2(., Time, weight)) %>% 
        ggplot(aes(Time, weight, colour = Chick)) + geom_point() + geom_line() + 
        geom_vline(xintercept = 21, alpha = 0.2) + labs(title = 'reorder')
)
```
