# R packages: Zhuoer's development style



see https://github.com/dongzhuoer/minir

I try to minimize (no function, no data), but still overwhelm for beginners.

`zzz.R` suggest by _R packages_, `aaa.R` not specail meaning (not need to be first file), just contrast `zzz.R`.


## basics



### dependency

- All needed packages should be listed in `Imports` (we don't use `Depends`).
- `Remotes` only specifies where to _find_ a package _listed_ in `Imports`, if not from CRAN.



### data

For example of using cache, refer to [rGEO.data](https://github.com/dongzhuoer/rGEO.data/blob/master/R-raw/data.Rmd)

- `devtools::ues_data()` store each object and its name in separate `.rds` file under `data/` ^[Although the `.rds` file is usually named after the corresponding object, it's not compulsory. You can change the filename and rebuild the package, the name of exported data remain unchaged.].
- `devtools::ues_data(internal=TURE)` store all the internal data in a single file, `R/sysdata.rda`. Every time you call this function, the old file would be overwrited.



### testthat

If you want to run arbitrary code, you need add a test, otherwise it would be skipped.

```r
testthat::test_that('foobar()', {
    testthat::expect_true(T);

    print('haha, I\'m not skipped')    
});
```



### Travis CI

I assume that all my packages use testthat and contain at least a `test-*.R` ^[`devtools::test()` only requires `tests/testthat` directory, but `testthat::test_dir()` dictate there must be a R script].

You can use `tolower(Sys.getenv('CI')) == 'true')` to selectively run some tests on Travis.






## document



### function documentation

```r
#' @title A sentence in one line.
#'
#' @description A paragraph which briefly describe what the function does,
#'   usually supplements the title.
#'
#' @param name class. One sentence or paragraph.
#' @param name1 integer scalar. When the length must be 1, add 'scalar'.
#' @param name2 string. Means _character scalar_, consistent with C++.
#' @param ...  passed onto [foo::bar()].
#'
#' @return list
#'
#'   1. x: string. name.
#'
#'   1. y: numeric. height.
#'
#' @details A (often long) section which goes into details about how the
#'   function works.
#'
#'   Many paragraphs is ok.
#'
#' @seealso [foobar2]
#'
#' @examples
#' foobar(1:5)
#'
#' \donotrun {
#'     foobar("You're a doubi.")
#' }
#'
#' @section Other points: One or many paragraphs.
#'
#' @export
foobar <- function(...) {
}
```

- If the documentation doesn't update after `Install and Restart`, you should go to **Build** panel => `More` => `Configure Build Tools...` => `Configure...`, and check **Build & Reload** ^[The option is saved in `.Rproj.user/***/build_options`, `***` is a hash string like `6AA761D1` to distinguish users.]. 
- In `@examples`, use `foobar()` rather than `pkg::foobar()`. Combined with `More` => `Load All`, it makes testing examples very convenient ^[And it's a convention to assume that user have `library()` your package before running examples.]. 
- multiple functions [example](https://github.com/dongzhuoer/rexseek/blob/93f6e0c41f3362f2612778d985450fb687188041/R/normalization.R#L91)
- The last line controls the accessibility:
  1. `@export`, full support.
  1. `@keywords internal`, can use `?` to get help, but not listed in index.
  1. `@noRd`, documetation not available.



### pkgdown

- [edit](https://pkgdown.r-lib.org/reference/build_site.html#yaml-config) `pkgdown/_pkgdown.yml` [^I used to put it `inst/`, but I realized I'm not interested in building site from CRAN source package
]
- `pkgdown/extra.css` and `pkgdown/extra.js` allows minor customisation 
- build into `pkgdown/output` since pkgdown refuses to overwrite non-empty directory (`pkgdown/`)




### vignette

In a vignette, you should show things happened _after_ your package is installed, which means `library(pkg)` and `system.file(package = 'pkg')` ^[This is a little counter-intuitive when you are developing a package, since you need to edit source code, `Install and Restart`, edit and build vignette, and `Install and Restart` again to see the result. However, it makes the user feel quite natural.].






## namespace

Usually, `@importFrom` should appear at the function using it (multiple times is okay, roxygen2 handles that).

But some (_too_) common things are put into `R/aaa.R`:

```r
#' @importFrom magrittr %>%
#' @export
magrittr::`%>%`
```

1. Here we import `%>%`, since `devtools::test()` won't help us, even if we use `%>%` in our function without importing it. Actually, when tests are executed, **testthat** is loaded, even worse, it exports `%>%` (see `` testthat::`%>%` `` & `` ?`%>%` ``).
2. Exporting `%>%` makes it easier to write tests & examples. Althougth you can live without it ^[Again, **testthat** exports %>%. Even if you use `%>%` in tests without exporting it, no one warns you.], users would be confusing when run the examples, since the basic assumption is that they have  `library()` your package.

Opionally, you can also import other pipes:

```r
#' @importFrom magrittr %T>%
#'
NULL
```

> `NULL` is necessary, otherwise the command would be ignored.






## developing



### IDE

You can use [RStudio](https://www.rstudio.com/products/RStudio/) or [VSCode](https://code.visualstudio.com/) (actually I use both).

I have included a few convenient tasks for VSCode.



### maintaining

- Not Yet Used
  ```{r, error = TRUE, collapse = TRUE}
  hypotenuse <- function(x, y, p = 2) {
    if (!missing(p))    .NotYetUsed("p")
    sqrt(x ^ 2 + y ^ 2)
  }
  
  hypotenuse(5, 12)
  hypotenuse(5, 12, 1)   
  ```

- Not Yet Implemented
  ```{r, error = TRUE, collapse = TRUE}
  triangular <- function(n) {
    .NotYetImplemented()
  }
  
  triangular()
  ```

- Deprecated
  ```{r, error = TRUE, collapse = TRUE}
  hypotenuse <- function(x, y, p = 2) {  
    .Deprecated("p_norm")
    (x ^ p + y ^ p) ^ (1 / p)
  }
  
  hypotenuse(5, 12) 
  ```

- Defunct
  ```{r, error = TRUE, collapse = TRUE}
  hypotenuse <- function(x, y, p = 2) {  
    .Defunct("p_norm")
  }
  
  hypotenuse(5, 12) 
  ```






## Rcpp {#rcpp-in-package} 



### set up

1. create a package (copy `minir`)
1. `usethis::use_rcpp()`
1. create a `.cpp` file in `src/`
1. set up appropriate ^[You should make it as portable as possible. For example, `CXX = clang++` is a bad idea here, since others may haven't install it and your code shouldn't depend on certain complier] `src/Makevars`, like: 
   ```
   PKG_CXXFLAGS = -std=c++17
   ```
1. add to `zzz.R`
   ```r
   .onUnload <- function (libpath) {
     library.dynam.unload("mypackage", libpath)
   }
   ```



### exporting & importing C++ code

refer to [here](https://bookdown.dongzhuoer.com/zhuoer/r-pkgs-revised/cpp.html#cpp-export)

If you want to export some header files, put them into `inst/include`, [example](https://github.com/dongzhuoer/Rcppzhuoer/tree/master/inst/include)

In rapidly development, you may add or remove `// [[Rcpp::interfaces(r, cpp)]]` in a `.cpp`, then you should use `Load All` since `Build` or `Document` would fail.



### best practice

1. use `Build` or `Load All` intead of sourcing individual `.cpp` file. ^[Actually, writing a pakcage is the standard solution when you want to use more than one `.cpp` file.]
1. `.cpp` file in a package shouldn't contain `// [[Rcpp::plugins(cpp17)]]`, that's the job of `src/Makevars`
