```{r include=FALSE, cache=FALSE}
# adapted from https://github.com/hadley/r4ds/blob/master/_common.R

set.seed(1014)
options(digits = 3)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  error = TRUE,
  cache = TRUE,
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  out.width = "70%"
)

options(dplyr.print_max = 10)

```
# plot with R






## ggplot2

- `labs(x = NULL)` can suppress x axis label

- color

```{r color}
scales::show_col(scales::hue_pal()(4))
```

- legend

```{r}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(aes(colour = factor(cyl), shape = factor(vs))) +
    labs(x = "Weight (1000 lbs)", y = "Fuel economy (mpg)", colour = "Cylinders", shape = "Transmission") + 
    theme(legend.background = element_rect(colour = 'purple', fill = 'pink', size = 3, linetype='dashed'), 
		  legend.margin = margin(20,15,10,5), 
		  legend.box.background = element_rect(colour = 'blue', fill = 'yellow', size = 5, linetype='solid'), 
		  legend.box.margin = margin(10,20,30,40))
```






## multi-plot

we use `cowplot::plot_grid()` as reference

- for **gtable**, `rbind()` & `cbind()` just align plots, and you have to use `ggplotGrob()` to convert ggplot object, [example](https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html#cb5)
- for **egg**, `ggarrange()` doesn't provide more feature, but some figure explains how to align panels: [one](https://github.com/baptiste/egg#aligning-complex-ggplots) & [another](https://github.com/jwdink/egg#aligning-complex-ggplots) ^[there are two GitHub repo, the former releases to CRAN]
- for **gridExtra**, `grid.arrange()` can't add label, and syntax for layout is more complex. But can add title, contain hole, nest and multi-page [vignette](https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html)

- **patchwork** provides a nice `+` syntax, but it remains to see whether I have to `library()` it, [home](https://github.com/thomasp85/patchwork)

- **multipanelfigure** 's syntax is very different, but also very versatile, the author even use it to produce a supp figure of a biomedical paper [guide](file:///home/zhuoer/Downloads/v84c03.pdf)

Thanks to this [vignette](https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html), which contains many other useful tips:

- Plot insets
- share legend 
- convert table to plot

following are R code I write when reproducing that vignette

```r
gl <- lapply(1:4, function(ii) 
    grid::grobTree(grid::rectGrob(gp=grid::gpar(fill=ii, alpha=0.5)), grid::textGrob(paste0('plot', ii)))) 

grid.arrange(
    grobs = gl,
    widths = c(2, 1, 1),
    layout_matrix = rbind(c(1, 2, NA),
                          c(3, 3, 4))
)

p1 <- qplot(mpg, wt, data = mtcars, colour = cyl)
p2 <- qplot(mpg, data = mtcars) + ggtitle("title")
p3 <- qplot(mpg, data = mtcars, geom = "dotplot")
p4 <-
    p1 + facet_wrap( ~ carb, nrow = 1) + theme(legend.position = "none") +
    ggtitle("facetted plot")
grid.arrange(
    p1, p2, p3, p4,
    widths = c(1, 1, 1),
    layout_matrix = rbind(c(1, 2, 3),
                          c(4, 4, 4))
)


g <- ggplotGrob(qplot(1, 1) +
                    theme(plot.background = element_rect(colour = "black")))
qplot(1:10, 1:10) +
    annotation_custom(
        grob = g,
        xmin = 1,
        xmax = 5,
        ymin = 5,
        ymax = 10
    ) +
    annotation_custom(
        grob = grid::rectGrob(gp = grid::gpar(fill = "white")),
        xmin = 7.5,
        xmax = Inf,
        ymin = -Inf,
        ymax = 5
    )
```






## graphics

```{r}
library(nutshell)
library(learningr)

as_ggplot_list <- function(...) {
    rlang::enexprs(...) %>% purrr::map(as.expression) %>% purrr::map(ggplotify::as.ggplot)
}

data(doctorates, batting.2008, turkey.price.ts, toxins.and.cancer, yosemite)
bat = plyr::mutate(
    batting.2008, 
    PA = AB + BB + HBP + SF + SH,
    OBP = (H + BB + HBP)/(AB + BB + HBP + SF)
)
bat.w.names <- plyr::mutate(
    bat, 
    throws = as.factor(throws), 
    bats = as.factor(bats), 
    AVG = H/AB
)
toc = toxins.and.cancer

data(obama_vs_mccain, crab_tag)
obama_vs_mccain = obama_vs_mccain[!is.na(obama_vs_mccain$Turnout), ]
ovm = within(obama_vs_mccain, Region <- reorder(Region, Obama, median))
```



### univariate plot

#### pie chart

```{r fig.dim=c(6,4)}
c(
    "Fresh and forzen" = 7752,
    "Reduced to meal, etc." = 1166,
    "Canned" = 463,
    "Cured" = 108
) %>% pie(main = "domestic catch 2006")
```

#### bar chart

transform data.frame into a matrix for barplot

```{r}
doctorates.m <- as.matrix(doctorates[,-1])
rownames(doctorates.m) <- doctorates[,1]
doctorates.m
```

If the vector or matrix isn't named, you need to use `names.arg` to specify the labels.

```{r fig.asp=0.75}
barplot(doctorates.m[1,])
barplot(doctorates.m, beside = T, horiz = T, legend = T)
## we set `ylim` to make room for the legend
barplot(
    t(doctorates.m), legend = T, ylim = c(0,66000), 
    args.legend = list(ncol = 2)
)
```

#### histogram

```{r fig.dim=c(8,6)}
as_ggplot_list(
    hist(obama_vs_mccain$Obama, main = 'number'),
    hist(obama_vs_mccain$Obama, freq = F, main = 'frequency')
) %>% gridExtra::grid.arrange(grobs = ., nrow = 1)
```

how to specify binwidth (`breaks` parameter)

```{r}
hist(obama_vs_mccain$Obama, 4, main = "An exact number of bins")
hist(obama_vs_mccain$Obama, seq.int(0, 100, 5), main = "A vector of bin edges")
hist(obama_vs_mccain$Obama, nclass.scott, main = "A function for the number of bins")
# you can also use the name of a method, such as `"FD"`
hist(obama_vs_mccain$Obama, function(x) {seq(min(x), max(x), length.out = 50)}, main = "A function for the bin edges")
```

#### denstity plot

```{r}
plot(density(bat[bat$PA > 25,"PA"]))
rug(bat[bat$PA > 25,"PA"])  #added to the above
```

#### quantile plot

```{r}
qqnorm(bat$AB)
```

#### time series plot

`acf()` computes and plots the autocorrelation function for a time series:

```{r fig.asp=0.8, out.width='50%'}
as_ggplot_list(
    plot(turkey.price.ts),
    acf(turkey.price.ts)
) %>% gridExtra::grid.arrange(grobs = ., nrow = 1)
```



### bivariate plot

#### scatter plot

```{r fig.width=8}
as_ggplot_list(
    plot(obama_vs_mccain$Income, obama_vs_mccain$Turnout, col = "violet", pch = 20),
    plot(obama_vs_mccain$Income, obama_vs_mccain$Turnout, log = "xy")
) %>% gridExtra::grid.arrange(grobs = ., nrow = 1)
```

pairs of correlation

```{r fig.width=9}
plot(bat[,c("H","R","SO","BB","HR")], main = '')
```


add text labels

```{r fig.width=8}
plot(
    toc$air_on_site/toc$Surface_Area, toc$deaths_lung/toc$Population,
    xlab = "Air Realse Rate of Toxic Chemicals", ylab = "Lung Cancer Death Rate"
)
text(
    toc$air_on_site/toc$Surface_Area, toc$deaths_lung/toc$Population, 
    cex = 0.5, labels = toc$State_Abbrev, adj = c(0, -1)
)
```

point shape is specified by `pch` (plotting character)

```{r}
plot(1:25, pch = 1:25, bg = "blue")
```


#### line plot

line plots are created in the same way as scatterplots, except that they take the argument **type = "l"**

```{r fig.keep='all'}
plot(crab_tag$daylog$Date, -crab_tag$daylog$Max.Depth, type = "l", ylim = c(-max(crab_tag$daylog$Max.Depth), 0))
## use `lines()` to draw additional lines on an existing plot
lines(crab_tag$daylog$Date, -crab_tag$daylog$Min.Depth, col = "blue")
```

#### box plot

1. The box shows the interquartile range, which contains values between the 25th and 75th percentile;  
2. The line inside the box shows the median;
3. The two **whiskers** on either side of the box show the adjacent values, which are intended to show extreme values;  
4. When there are values far outside the range we would expect for normally distributed data, those outlying values are plotted separately. 
5. Specifically, here is how the adjacent values are calculated:  
   - the upper adjacent value is the value of the largest observation that is less
than or equal to the upper quartile plus 1.5 times the length of the interquartile
range; 
   - the lower adjacent value is the value of the smallest observation that is greater
than or equal to the lower quartile less 1.5 times the length of the interquartile range.

```{r fig.width=10}
as_ggplot_list(
    boxplot(Obama ~ Region, data = obama_vs_mccain),
    boxplot(Obama ~ Region, within(obama_vs_mccain, Region <- reorder(Region, Obama, median)))
) %>% gridExtra::grid.arrange(grobs = ., nrow = 1)
```

#### plot categorical data

```{r}
# conditional density of a set of categories dependent on a numeric value
cdplot(bats ~ AVG, bat.w.names, subset = (bat.w.names$AB > 100))
spineplot(bats~throws, bat.w.names)
mosaicplot(bats~throws, bat.w.names, color = TRUE)
assocplot(table(bat.w.names$bats, bat.w.names$throws), xlab = "Throws", ylab = "Bats")
```



### trivariate plot

we need to make two transformations:

1. In the data file, values move east to west (or left to right) as x indices increase and from north to south (or top to bottom) as y indices increase. Unfortunately, **persp**  plots increasing y coordinates from bottom to top.  
2. We need to select only a square subset of the elevation points.

To plot the figure, we rotate the image by 225?? (through **theta=225**), change the viewing angle to 20?? (**phi=20**), adjuste the light source to be from a 45?? angle (**ltheta=45**) and set the shading factor to 0.75 (**shade=.75**) to exaggerate topological features.

```{r}
row = nrow(yosemite)
col = ncol(yosemite)
halfdome <- yosemite[(row - col + 1):row, col:1]
```

#### `persp()` and `headmap()`



```{r}
# three-dimensional surface for a specific perspective.
persp(halfdome, col = grey(.25), border = NA, expand = .15, theta = 225, phi = 20, ltheta = 45, lphi = 20, shade = .75)
```

```{r}
heatmap(yosemite)
```

#### `image()` and `contour()`

`image()` plot a matrix of data points as a grid of boxes, color coding the boxes based on the intensity at each location.

- `asp` specifies an aspect ratio that matches the dimensions of the data;  
- `ylim` specifies that data is plotted from top to bottom

```{r}
image(yosemite, asp = col/row, col = sapply((0:32)/32, gray))
contour(yosemite, asp = col/row, ylim = c(1,0))
```

Contours are commonly added to existing image plots:

```{r}
image(yosemite, asp = col/row, ylim = c(1,0), main = 'image + contour')
contour(yosemite, asp = col/row, ylim = c(1,0), add = T)
```



### customizing charts

#### graphical parameters

- get

```{r}
par("bg")
head(par(),3)
```

- set

```{eval = F}
par(bg = "transparent")
```

#### mutliple plots

```{r fig.dim=c(6, 5)}
par(mfcol = c(2,2))

pie(5:1)
plot(1:400,rnorm(400))
qqnorm(rnorm(100))
hist(rnorm(10000))
```

#### low-level graphics functions

- `abline()`

```{r}
plot(x = c(0,10),y = c(0,10), main = 'add lines')
abline(h = 4.5, col = 'red')
abline(v = 3.5, col = 'blue')
abline(a = 1, b = 1, col = 'green')
abline(coef  =  c(10,-1))
abline(h = 1:10,v = 1:10, col = 'gray')
```

- `polygon()`

```{r}
plot(x = c(0,10),y = c(0,10), main = 'add polygon')
polygon(x = c(2,2,4,4),y = c(2,4,4,2))
```






## lattice

```{r}
library(lattice)
library(nutshell)
library(learningr)

data(obama_vs_mccain, crab_tag)
data(births2006.smpl, tires.sus, sanfrancisco.home.sales, team.batting.00to08)

births = births2006.smpl
san = sanfrancisco.home.sales
obama_vs_mccain = obama_vs_mccain[!is.na(obama_vs_mccain$Turnout), ]
ovm = within(obama_vs_mccain, Region <- reorder(Region, Obama, median))
ovm2 <- ovm[!(ovm$State %in% c("Alaska", "Hawaii")), ]
```

### introduction

#### how lattice works

1. The end user calls a high-level lattice plotting function.
2. The **lattice** function examines the calling arguments and default parameters, assembles a lattice object, and returns the object. (Note that the class of the object is actually **trellis**.)
3. The user calls **plot.trellis** or **plot.trellis** with the lattice object as an argument. (This typically happens automatically on the R console.)
4. The function **plot.trellis** sets up the matrix of panels, assigns packets to different panels and then calls the panel function specified in the lattice object to draw the individual panels.

#### a simple example

```{r}
(d <- data.frame(x = c(0:9), y = c(1:10), z = c(rep(c("a", "b"), times = 5))))
```

```{r fig.dim=c(4, 2.5)}
xyplot(y~x, data = d)
xyplot(y~x, groups = z, data = d)
xyplot(y~x | z, data = d)
xyplot(
    y~x | z, data = d, 
    panel = function(...){
        panel.abline(a = 1,b = 1)
        panel.xyplot(...)
    }
)
```

#### store lattice plots and update

```{r}
(lat1 <- xyplot(Turnout ~ Income | Region,obama_vs_mccain))
(lat2 <- update(lat1, col = "violet", pch = 20))
```



### univariate plot

#### bar chart

- one-dimentional table

```{r}
(births.dow <- table(births$DOB_WK))

barchart(births.dow)
```

- two-dimensional table

```{r}
births2006.dm <- plyr::mutate(births[births$DMETH_REC != "Unknown",],
                              DMETH_REC = as.factor(as.character(DMETH_REC)))
(dob.dm.tbl <- with(births2006.dm, table(WK = DOB_WK,MM = DMETH_REC)))
```

```{r}
barchart(dob.dm.tbl)
barchart(dob.dm.tbl, stack = F, auto.key = T)
barchart(dob.dm.tbl, horizontal = F, groups = F)
```


- data frame

```{r}
ovm2 %>% tibble::as_tibble()
```

```{r}
barchart(State ~ Catholic, ovm2)
barchart(State ~ Catholic + Protestant + Non.religious + Other, ovm2, stack = T)
```


#### dot plot

- three-dimensional table

```{r}
(dob.dm.tbl.alt <- with(births2006.dm, table(week = DOB_WK, month = DOB_MM, meths = DMETH_REC)))

dotplot(dob.dm.tbl.alt, groups = T, auto.key = T, stack = F)
```

- data frame

```{r}
tires.sus$Speed_At_Failure_km_h %>% tibble::as_tibble()

dotplot(as.factor(Speed_At_Failure_km_h) ~ Time_To_Failure | Tire_Type, tires.sus)
```

#### histogram

```{r}
histogram(~ Obama, obama_vs_mccain)    
histogram(~ Obama, obama_vs_mccain, breaks = 10)
histogram(~ Obama, obama_vs_mccain, type = "percent")
histogram(~ Obama, obama_vs_mccain, type = "count")
histogram(~ Obama, obama_vs_mccain, type = "density")
histogram(~DBWT | DPLURAL, births)
histogram(~DBWT | DPLURAL, births, layout = c(1,5))
```

#### denstity plot

```{r}
(plur = unique(as.character(births$DPLURAL))[4:5]) #quadruplet or higher
```

```{r}
#By default, densityplot will draw a strip chart under each chart:
densityplot(~DBWT | DPLURAL, births, subset = (DPLURAL %in% plur))
#turn off the strip by plot.points = F because the data set is so big
densityplot(~DBWT | DPLURAL, births, layout = c(1,5), plot.points = F)
densityplot(~DBWT, groups = DPLURAL, births, plot.points = F, auto.key = T)
```

#### strip plot

You can think of strip plots as one-dimensional scatter plots:

```{r}
#jitter.data=T adds some random vertical noise to make the points easier to read
stripplot(~DBWT, births, jitter.data = T, subset = (DPLURAL %in% plur))
```

#### quantile plot

```{r}
qqmath(rnorm(100000))
qqmath(~DBWT | DPLURAL, births, pch = 19, cex = .25)
qqmath(~price, san)
qqmath(~log(price), san)
qqmath(~log(price), groups = bedrooms, auto.key = T, type = "smooth",
       subset(san, !is.na(bedrooms) & bedrooms > 0 & bedrooms < 7))
#using subset parameter causes unused factor levels: 
qqmath(~log(price), groups = bedrooms, san, auto.key = T, type = "smooth",
       subset = !is.na(bedrooms) & bedrooms > 0 & bedrooms < 7)
#we use subset parameter because NA is cleaned up:
qqmath(~log(price), groups = Hmisc::cut2(squarefeet, g = 6), san,
       subset = !is.na(squarefeet), type = "smooth", auto.key = T)
```


### bivariate plot

#### scatter plot

```{r}
xyplot(Turnout ~ Income, obama_vs_mccain, col = "violet", pch = 20)
## use `list(log = TRUE)` to log-scale both axes
xyplot(Turnout ~ Income, obama_vs_mccain, scales = list(y = list(log = TRUE))) 
#relation = "same" means that each panel shares the same axes.
#alternating = T(default) means axis ticks for each panel are drawn on alternating sides of the plot, otherwise just the left and bottom 
xyplot(Turnout ~ Income | Region, obama_vs_mccain, layout = c(5, 2),
       scales = list(log = T, relation = "same", alternating = F))
```


```{r}
dollars.per.squarefoot <- with(san, mean(price / squarefeet, na.rm = T))
zips = subset(san, !is.na(squarefeet), zip) %>% table() %>% sort() %>% head(4*5) %>% names() %>% as.integer()
san.subset <- subset(san, zip %in% zips & price < 4000000 & squarefeet < 6000)

trellis.par.set(fontsize = list(text = 7))
xyplot(price~squarefeet | neighborhood, san.subset, pch = 19, cex = .2,
       strip = strip.custom(strip.levels = T, horizontal = T,
                            par.strip.text = list(cex = .8)),
       panel = function(...) {
           panel.abline(a = 0, b = dollars.per.squarefoot)
           panel.xyplot(...)
        }
)
```

### line plot

```{r}
xyplot(-Min.Depth + -Max.Depth ~ Date, crab_tag$daylog, type = "l")
```

### box plot

"bw" is short for "b (box) and w (whisker)",

```{r}
table(cut(san$date,"month"))

bwplot(log(price) ~ cut(date,"month"), san, scales = list(x = list(rot = 90)))
```


### trivariate plot

```{r}
levelplot(price~latitude + longitude, san)
levelplot(with(san, table(cut(longitude,40), cut(latitude,40))),
          scales = list(x = list(rot = 90, cex = .5), y = list(cex = .5)))
levelplot(
    with(san, tapply(price,list(cut(longitude, 40), cut(latitude, 40)), mean)),
    scales = list(draw = F))
levelplot(with(san, table(cut(longitude, 25),
                          cut(latitude, 25),
                          ifelse(bedrooms < 5, bedrooms, 5))),
          scales = list(draw = F))
```

- `lattice::contourplot()`
- `lattice::cloud()`
- `lattice::wireframe()`


### other plot

#### residual and fit-spread (RFS) plot
```{r}
model = lm(runs~singles + doubles + triples + stolenbases + homeruns + walks 
           + hitbypitch + sacrificeflies + caughtstealing, team.batting.00to08)
```

```{r}
rfs(model)
rfs(model,distribution  =  qnorm)
```


### lattice parameters

#### get
```{r}
trellis.par.get("clip")

head(trellis.par.get(), 3)
```

#### set
```{r}
trellis.par.set(list(axis.text = list(cex = .5)))
```

#### nicer show

```{r}
show.settings()
```

#### deep diving

There are a total of 378 parameters. However, there are only 46 unique parameters within these groups:
```{r}
length(unlist(trellis.par.get()))

(function() {
    n = names(trellis.par.get())
    p = NA
    for (i in 1:34) p = c(p,names(trellis.par.get(n[i])))
    length(table(p))
})()
```

