# yaml



This post mainly covers [yaml in R](https://github.com/viking/r-yaml), see [中文教程](http://www.ruanyifeng.com/blog/2016/07/yaml.html) for more information.


## yes or no, that is a question

```yaml
yes: yes
no: no
```

means

```{r echo=FALSE, comment=NA}
yaml::yaml.load('yes: yes\nno: no')
```

You should add quotes to avoid being interpreted as `T` or `F`

```yaml
'yes': yes
'no': 'no'
```

means

```{r echo=FALSE, comment=NA}
yaml::yaml.load('"yes": yes\n"no": "no"')
```

There's actually a [long list of reserved words](http://yaml.org/type/bool.html) with this behavior:

```
y|Y|yes|Yes|YES|n|N|no|No|NO
|true|True|TRUE|false|False|FALSE
|on|On|ON|off|Off|OFF
```

Wa, the world is so nice!



## vector

There are two ways to represent vectors:

```yaml
rmd_files: ["index.Rmd", "body.Rmd"]
```

```yaml
rmd_files:
  - "index.Rmd"
  - "body.Rmd"
```

the following ways is also okay, but not recommended (at least by me)

```yaml
rmd_files: [
  "index.Rmd", 
  "body.Rmd"
]
```

or even

```yaml
rmd_files: [
  "index.Rmd", 
  "body.Rmd",
]
```



## list

缩进的空格数目不重要，只要相同层级的元素左侧对齐即可

Values can also be lists of values. You just need to indent the values by two more spaces, e.g.,

```yaml
output:
  bookdown::gitbook:
    split_by: "section"
    split_bib: no
```

It is a common mistake to forget to indent the values. For example, the following data

```yaml
output:
html_document:
toc: yes
```

actually means

```yaml
output: null
html_document: null
toc: yes
```

instead of what you probably would have expected:

```yaml
output:
  html_document:
    toc: yes
```
