# YAML, etc

This chapter aims to include simple language, althought currently only include YAML






## YAML

> 用相同数目的空格（我推荐 4个）缩进元素，来表示层级关系。

See [中文教程](http://www.ruanyifeng.com/blog/2016/07/yaml.html) for detailed guide. Here we mainly talked about YAML used for configuration.



### basics

```yaml
age: 27
men: [John, Bill]
women:
  - Mary
  - Susan
websites:
  YAML: yaml.org 
```



### yes or no, that is a question

这一节是我无意中发现的，当时在纠结到底应该写 `yes|no` 还是 `true|false`（最后我选择了前者，因为 YAML 强调 human-readable）。

```yaml
yes: yes
no: no
```

means

```{r echo=FALSE, comment=NA}
yaml::yaml.load('yes: yes\nno: no')
```

You should add quotes to avoid being interpreted as `T` or `F`

```yaml
'yes': yes
'no': 'no'
```

means

```{r echo=FALSE, comment=NA}
yaml::yaml.load('"yes": yes\n"no": "no"')
```

There's actually a [long list of reserved words](http://yaml.org/type/bool.html) with this behavior:

```
y|Y|yes|Yes|YES|n|N|no|No|NO
|true|True|TRUE|false|False|FALSE
|on|On|ON|off|Off|OFF
```

Wa, the world is so nice!






## TOML

To be written.
